---
output:
  word_document: default
  html_document: default
---
```{r, include = FALSE}

library(tidyverse)
library(tidymodels)
library(rpart)
library(rpart.plot)
library(RColorBrewer) 
library(rattle)
library(caret)

parole <- read_csv("parole.csv")

parole <- parole %>% mutate(male = as_factor(male)) %>%
  mutate(male = fct_recode(male, "female" = "0", "male" = "1")) %>%
  mutate(race = as_factor(race)) %>%
  mutate(race = fct_recode(race, "otherwise" = "2", "white" = "1")) %>%
  mutate(state = as_factor(state)) %>%
  mutate(state = fct_recode(state, "Kentucky" = "2", "other" = "1", "Virginia" = "4", "Louisiana" = "3")) %>%
  mutate(crime = as_factor(crime)) %>%
  mutate(crime = fct_recode(crime, "larceny" = "2", "other" = "1", "driving related crime" = "4", "drug related crime" = "3")) %>%
  mutate(multiple.offenses = as_factor(multiple.offenses)) %>%
  mutate(multiple.offenses = fct_recode(multiple.offenses, "multiple offenses" = "1", "otherwise" = "0")) %>%
  mutate(violator = as_factor(violator)) %>%
  mutate(violator = fct_recode(violator, "violated parole" = "1", "parole complete without violation" = "0")) 

```


```{r Task 1}
#Task 1
set.seed(12345)
parole_split <- initial_split(parole, prop = 0.70, strata = violator)
train <- training(parole_split)
test <- testing(parole_split)


```

```{r Task 2}
#Task 2
parole_recipe = recipe(violator ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree() %>% 
  set_engine("rpart", model = TRUE) %>%
  set_mode("classification")

parole_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(parole_recipe)

parole_fit = fit(parole_wflow, train)
```

```{r}
tree = parole_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")
#rpart.plot(tree, tweak =1.5)
fancyRpartPlot(tree, tweak=1.4)
```

Task 3
A parolee from Louisana would 59% likelihood to complete parole without violation.  The parole who served 5 years would be 57% likely to complete parole without incident while if over 40 years old be 46% likely to violate parole.  Multiple offenses were not a significant factor for a Louisiana parolee in this tree.


```{r Task 4}
#Task 4
parole_fit$fit$fit$fit$cptable
```

The xerror corresponding to the optimal cp is 1.314815 and is associated with 8 splits.  Yes, the tree in Task 2 is associated with this optimal cp.

```{r Task 5}
#Task 5
set.seed(123)
folds = vfold_cv(train, v = 5)

parole2_recipe = recipe(violator ~., train) %>%
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid = grid_regular(cost_complexity(),
                          levels = 25) 

parole2_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(parole2_recipe)

tree_res = 
  parole2_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res

```

```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2)
```

```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

```{r}
final_wf = 
  parole2_wflow %>% 
  finalize_workflow(best_tree)
```

```{r}
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

#fancyRpartPlot(tree, tweak = 1.5) 
```

```{r}
treepred = predict(final_fit, train, type = "class")
head(treepred)
```

```{r}
confusionMatrix(treepred$.pred_class,train$violator,positive="parole complete without violation")
```

```{r}
treepred_test = predict(final_fit, test, type = "class")
head(treepred_test)
```

```{r}
confusionMatrix(treepred_test$.pred_class,test$violator,positive="parole complete without violation")
```

```{r}
parole3_recipe = recipe(violator ~., train) %>% 
  step_dummy(all_nominal(),-all_outcomes())

tree_model = decision_tree(cost_complexity = tune()) %>% 
  set_engine("rpart", model = TRUE) %>% 
  set_mode("classification")

tree_grid = expand.grid(cost_complexity = seq(0.001,0.01,by=0.001))

parole3_wflow = 
  workflow() %>% 
  add_model(tree_model) %>% 
  add_recipe(parole3_recipe)

tree_res = 
  parole3_wflow %>% 
  tune_grid(
    resamples = folds,
    grid = tree_grid
    )

tree_res
```

```{r}
tree_res %>%
  collect_metrics() %>%
  ggplot(aes(cost_complexity, mean)) +
  geom_line(size = 1.5, alpha = 0.6) +
  geom_point(size = 2) +
  facet_wrap(~ .metric, scales = "free", nrow = 2) 
```

```{r}
best_tree = tree_res %>%
  select_best("accuracy")

best_tree
```

```{r}
final_wf = 
  parole3_wflow %>% 
  finalize_workflow(best_tree)
```

```{r Task 6 and 7}
#Task 6 and 7 
final_fit = fit(final_wf, train)

tree = final_fit %>% 
  pull_workflow_fit() %>% 
  pluck("fit")

final_fit$fit$fit$fit$cptable

fancyRpartPlot(tree, tweak = 1.5) 
```

```{r}
final_fit$fit$fit$fit$cptable
```

The optimal cp value is 8
