---
output:
  word_document: default
  html_document: default
---


```{r, include = FALSE}
#install.packages("factoextra")

library(tidymodels)
library(tidyverse)
library(cluster) #algorithms for clustering
library(factoextra) #visualization

trucks <- read_csv("trucks.csv")
```
```{r}
str(trucks)
summary(trucks)
```

```{r}
ggplot(trucks, aes(Distance, Speeding)) +
  geom_point(alpha = 0.4) + theme_bw()
```

Task 1 
There appears to four clusters and a correlation between distance traveled and speed.

```{r}
kmeans_recipe = recipe(~ Distance + Speeding, trucks) 

trucks2 = kmeans_recipe %>% 
  step_scale(all_numeric()) %>%
  step_center(all_numeric()) 

trucks2 <- prep(trucks2, trucks)

trucks_cleaned = bake(trucks2, trucks) 
```

Task 2 

```{r}
summary(trucks_cleaned)
summary(trucks)
```



```{r}
set.seed(64)
kclust = kmeans(trucks_cleaned, centers = 2)
kclust
```

```{r}

points <- augment(kclust, trucks) 
#points = points %>% select(-.cluster)
points
```


```{r}
ggplot(points, aes(Distance, Speeding, color = .cluster)) +
  geom_point(alpha = 0.4) + theme_bw()

```

Task 3
The kmeans clustings show two loose clusters.  It appeas more refinement may be needed or possibly use the ggplot(clusterings, aes(k, tot.withinss)) command to see if there's a more optimal k value.  




```{r}
set.seed(412)
clusts = 
  tibble(k = 1:8) %>% #try from 1 to 8 clusters
  mutate(
    kclust = map(k, ~kmeans(trucks, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, trucks)
  )

clusts
```

```{r}
clusters =
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
p1 = 
  ggplot(assignments, aes(x = Distance, y = Speeding)) +
  geom_point(aes(color = .cluster), alpha = 0.8) + 
  facet_wrap(~ k)
p1
```



```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

Task 4 
4 appears to be the best k value

```{r}
trucks_scaled = scale(trucks) 
summary(trucks_scaled)
```

```{r}
set.seed(2521)
clusts = 
  tibble(k = 1:10) %>%
  mutate(
    kclust = map(k, ~kmeans(points, .x)),
    tidied = map(kclust, tidy),
    glanced = map(kclust, glance),
    augmented = map(kclust, augment, points)
  )
clusts
```

```{r}
clusters =
  clusts %>%
  unnest(cols = c(tidied))

assignments = 
  clusts %>% 
  unnest(cols = c(augmented))

clusterings = 
  clusts %>%
  unnest(cols = c(glanced))
```

```{r}
ggplot(clusterings, aes(k, tot.withinss)) +
  geom_line() +
  geom_point() + theme_bw()
```

Task 5

3 or 4 clusters appear to be the appropriate k value 

```{r}
set.seed(64)
kclust = kmeans(trucks_cleaned, centers = 4)
kclust 
```


```{r}
points <- augment(kclust, trucks) 
points
```

```{r}
ggplot(points, aes(Distance, Speeding, color = .cluster)) +
  geom_point(alpha = 0.4) + theme_bw()
```

Task 6
A k value of 4 appears to be more optimal than the previous k value of 2.  The data visualization appears more meaningful to me.  