---
output:
  word_document: default
  html_document: default
---
```{r Libraries and Data }

library(tidyverse)
library(tidymodels)
library(e1071)
library(ROCR)
library(GGally)

parole <- read_csv("parole.csv")
parole <- parole %>% mutate(male = as_factor(male)) %>%
  mutate(male = fct_recode(male, "female" = "0", "male" = "1")) %>%
  mutate(race = as_factor(race)) %>%
  mutate(race = fct_recode(race, "otherwise" = "2", "white" = "1")) %>%
  mutate(state = as_factor(state)) %>%
  mutate(state = fct_recode(state, "Kentucky" = "2", "other" = "1", "Virginia" = "4", "Louisiana" = "3")) %>%
  mutate(crime = as_factor(crime)) %>%
  mutate(crime = fct_recode(crime, "larceny" = "2", "other" = "1", "driving related crime" = "4", "drug related crime" = "3")) %>%
  mutate(multiple.offenses = as_factor(multiple.offenses)) %>%
  mutate(multiple.offenses = fct_recode(multiple.offenses, "multiple offenses" = "1", "otherwise" = "0")) %>%
  mutate(violator = as_factor(violator)) %>%
  mutate(violator = fct_recode(violator, "violated parole" = "1", "parole complete without violation" = "0")) 


```

```{r Task 1}

set.seed(12345)
parole_split <- initial_split(parole, prop = 0.70, strata = violator)
train <- training(parole_split)
test <- testing(parole_split)


```

```{r Task 2 }

ggplot(parole,aes(x=male,fill = violator)) + geom_bar() + theme_bw()

```

```{r}
ggplot(parole,aes(x=race,fill = violator)) + geom_bar() + theme_bw()
```

```{r}
ggplot(parole,aes(x=multiple.offenses,fill = violator)) + geom_bar() + theme_bw()
```

```{r}
ggplot(parole,aes(x=crime,fill = violator)) + geom_bar() + theme_bw()
```

```{r}
ggplot(parole,aes(x=state,fill = violator)) + geom_bar() + theme_bw()
```

```{r  Task 2 continued}
#modl <- lm(violator ~ age + time.served + max.sentence, parole)
#modl
#modl1 <- parole  %>%
 # dplyr::select(age, time.served, max.sentence)
#ggpairs(modl1)

```
```{r}
t1 <- table(parole$violator, parole$max.sentence)
prop.table(t1, margin =2)
t1
```


I visualized several factors with bar graphs and visually identified gender, multiple offenders and state as significant factors in parole violation.  I thought a table based on max sentence length would be insightful; however, no meaningful insight was gained. 


```{r Task 3}
parole_recipe <- recipe(violator ~ multiple.offenses, parole) %>%
  step_dummy(all_nominal(), -all_outcomes())

parole_model <-
  logistic_reg() %>%
  set_engine("glm")

logreg_wflow <-
  workflow() %>%
  add_model(parole_model) %>%
  add_recipe(parole_recipe)

parole_fit <- fit(logreg_wflow, parole)
summary(parole_fit$fit$fit$fit)


```

As a single variable the module the AIC of 479.81 seems fine.   I consider the model on multiple offenses good as a test I modeled gender and it had a slightly higher AIC. 


```{r Task 3 option 2}
# parole2_recipe <- recipe(violator ~ male, parole) %>%
#   step_dummy(all_nominal(), -all_outcomes())
# 
# parole2_model <-
#   logistic_reg() %>%
#   set_engine("glm")
# 
# logreg2_wflow <-
#   workflow() %>%
#   add_model(parole2_model) %>%
#   add_recipe(parole2_recipe)
# 
# parole2_fit <- fit(logreg2_wflow, parole)
# summary(parole2_fit$fit$fit$fit)
```


```{r Task 4}
parolebest_recipe <- recipe(violator ~ multiple.offenses + male + state + crime + time.served, train) %>%
  step_dummy(all_nominal(), -all_outcomes())

parolebest_model <-
  logistic_reg() %>%
  set_engine("glm")

logregbest_wflow <-
  workflow() %>%
  add_model(parolebest_model) %>%
  add_recipe(parolebest_recipe)

parolebest_fit <- fit(logregbest_wflow, parole)
summary(parolebest_fit$fit$fit$fit)
```

The model has an AIC of 376.03.  At the very least the model is better than what I surmised was the best single factor of multiple offenses (AIC 487.17).  Multiple offenses and the states of Louisiana and Virginia are most significant.  I eliminated some factors to include in the model (e.g. race) due to their visual representations in the exploratory phase.  However, I suspected gender (male) especially, crime and time served would be significant model factors but were not so.  The lack of significance of those factors surprised me. 


```{r Task 5}

paroleT5_recipe <- recipe(violator ~ multiple.offenses + state + race, train) %>%
  step_dummy(all_nominal(), -all_outcomes())

paroleT5_model <-
  logistic_reg() %>%
  set_engine("glm")

logregT5_wflow <-
  workflow() %>%
  add_model(paroleT5_model) %>%
  add_recipe(paroleT5_recipe)

paroleT5_fit <- fit(logregT5_wflow, parole)
summary(paroleT5_fit$fit$fit$fit)


```

The AIC dropped and showed improvement over my "best" model!  Multiple offenses and Virginia were the most significant factors.  I am surprises race otherwise had any significance based on the ggplot visualizations. 




```{r Task 6}
parolee1 <- data.frame(state = "Louisiana", multiple.offenses = "multiple offenses", race = "white")
parolee2 <- data.frame(state = "Kentucky", multiple.offenses = "otherwise", race = "otherwise")
predict(paroleT5_fit, parolee1, type ="prob")

```

Parolee1 prediction

```{r}
predict(paroleT5_fit, parolee2, type ="prob")
```

Parolee2 prediction


```{r Task 7}

predictions = predict(paroleT5_fit, train, type = "prob") [2]
head(predictions)

```

```{r Task 7.1}
ROCRpred = prediction(predictions, train$violator)
ROCRperf = performance(ROCRpred, "tpr", "fpr")
plot(ROCRperf, colorize = TRUE, print.cutoffs.at=seq(0,1,by=0.1), text.adj=c(-0.2,1.7))

```


```{r Task 7.2}

as.numeric(performance(ROCRpred, "auc")@y.values)

```

```{r Task 7.3}
opt.cut = function(perf, pred){
  cut.ind = mapply(FUN=function(x, y, p){
    d = (x-0)^2 + (y-1)^2
    ind = which(d== min(d))
    c(sensitivity = y[[ind]], specificity = 1-x[[ind]],
      cutoff = p[[ind]])
  }, perf@x.values,perf@y.values, pred@cutoffs)
}
print(opt.cut(ROCRperf,ROCRpred))
```

```{r Task 7.4}

t4 = table(train$violator,predictions>0.1470858)
t4

```

```{r Task 7.5}
(t4[1,1]+t4[2,2])/nrow(train)

```


Task 8
Given the cutoff of 0.1470858 the accuracy is 0.8277228, sensitivity is 0.7586207, and specificity is 0.8210291. What are the implications of incorrectly classifying a parolee? The terms of the parolee’s parole may be too lenient. Oversight of a parolee may be less vigilant if a parolee is misclassified as not likely to violate parole.   

```{r Task 9}

t4 = table(train$violator,predictions>0.5)
t4
(t4[1,1]+t4[2,2])/nrow(train)
```

```{r Task 9.1}
t4 = table(train$violator,predictions>0.625)
t4
(t4[1,1]+t4[2,2])/nrow(train)
```

```{r Task 9.2}
t4 = table(train$violator,predictions>0.4)
t4
(t4[1,1]+t4[2,2])/nrow(train)
```


The probability threshold of >0.625 best maximizes accuracy on the training set.

```{r Task 10}
predictions = predict(paroleT5_fit, test, type = "prob") [2]
head(predictions)

t5 = table(test$violator,predictions>0.625)
t5
(t5[1,1]+t5[2,2])/nrow(test)
```

