---
output:
  word_document: default
  html_document: default
---
```{r, include = FALSE}

#install.packages(ranger)
#install.packages("randomForest")
#install.packages("titanic")
#install.packages("gridExtra")
#install.packages("vip")

library(titanic)
library(tidyverse)
library(tidymodels)
library(mice) #package for imputation
library(VIM) #visualizing missingness
library(ranger) #for random forests
library(randomForest) #also for random forests
library(caret)
library(naniar)
library(GGally)
library(gridExtra)
library(vip)

drug <- read_csv("drug_data-1.csv")

names(drug) = c("ID", "Age", "Gender", "Education", "Country", "Ethnicity",
"Nscore", "Escore", "Oscore", "Ascore", "Cscore", "Impulsive",
"SS", "Alcohol", "Amphet", "Amyl", "Benzos", "Caff", "Cannabis",
"Choc", "Coke", "Crack", "Ecstasy", "Heroin", "Ketamine", "Legalh",
"LSD", "Meth", "Mushrooms", "Nicotine", "Semer", "VSA")
str(drug)

drug[drug == "CL0"] = "No"
drug[drug == "CL1"] = "No"
drug[drug == "CL2"] = "Yes"
drug[drug == "CL3"] = "Yes"
drug[drug == "CL4"] = "Yes"
drug[drug == "CL5"] = "Yes"
drug[drug == "CL6"] = "Yes"

drug_clean = drug %>% mutate_at(vars(Age:Ethnicity), funs(as_factor)) %>%
  mutate(Age = factor(Age, labels = c("18_24", "25_34", "35_44","45_54", "55_64", "65_"))) %>%
  mutate(Gender = factor(Gender, labels = c("Male", "Female"))) %>%
  mutate(Education = factor(Education, labels =
                              c("Under16", "At16", "At17", "At18", "SomeCollege", "ProfessionalCert", "Bachelors", "Masters","Doctorate"))) %>%
  mutate(Country = factor(Country,labels = c("USA", "NewZealand", "Other", "Australia", "Ireland","Canada","UK"))) %>%
  mutate(Ethnicity = factor(Ethnicity,labels = c("Black", "Asian", "White", "White/Black", "Other","White/Asian", "Black/Asian"))) %>%
  mutate_at(vars(Alcohol:VSA), funs(as_factor)) %>%
  select(-ID)

str(drug_clean)

drug_clean = drug_clean %>% select(!(Alcohol:Mushrooms)) %>% select(!(Semer:VSA))
names(drug_clean)
```

```{r}

#Task 1
summary((drug_clean))
gg_miss_var(drug_clean)
vim_plot <- aggr(drug_clean, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.7)

```

```{r}

#Task 2
set.seed(1234)
drug_split <- initial_split(drug_clean, prop = 0.70, strata = Nicotine)
train <- training(drug_split)
test <- testing(drug_split)


```

```{r}
#Task 3
v1 <- ggplot(train, aes(x = Gender, fill = Nicotine)) + geom_bar(position = "fill")
v2 <- ggplot(train, aes(x = Age, fill = Nicotine)) + geom_bar(position = "fill")
v3 <- ggplot(train, aes(x = Education, fill = Nicotine)) + geom_bar(position = "fill")
v4 <- ggplot(train, aes(x = Country, fill = Nicotine)) + geom_bar(position = "fill")
grid.arrange(v1,v2,v3,v4)
```

```{r}
v5 <- ggplot(train, aes(x = Ethnicity, fill = Nicotine)) + geom_bar(position = "fill")
v6 <- ggplot(train, aes(x = Nscore, fill = Nicotine)) + geom_boxplot()
v7 <-  ggplot(train, aes(x = Escore, fill = Nicotine)) + geom_boxplot()
v8 <- ggplot(train, aes(x = Oscore, fill = Nicotine)) + geom_boxplot()
grid.arrange(v5,v6,v7,v8)
```


```{r}
v9 <- ggplot(train, aes(x = Ascore, fill = Nicotine)) + geom_boxplot()
v10 <- ggplot(train, aes(x = Cscore, fill = Nicotine)) + geom_boxplot()
v11 <- ggplot(train, aes(x = Impulsive, fill = Nicotine)) + geom_boxplot()
v12 <- ggplot(train, aes(x = SS, fill = Nicotine)) + geom_boxplot()
grid.arrange(v9,v10,v11,v12)
```

```{r}
#Task 4

set.seed(123)
rf_folds <- vfold_cv(train, v = 5)
```

```{r}
nicotine_recipe <- recipe(Nicotine ~., drug_clean) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model <- rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

nicotine_wflow <-
  workflow() %>% 
  add_model(rf_model) %>% 
  add_recipe(nicotine_recipe)

rf_grid <- grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)), 
  levels = 5
)

set.seed(123)
rf_res <- tune_grid(
  nicotine_wflow,
  resamples = rf_folds,
  grid = 10 
)
```

```{r}
rf_res %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

```{r}
#Task 5

nicotine_recipe_tr <- recipe(Nicotine ~., train) %>%
  step_dummy(all_nominal(), -all_outcomes())

rf_model_tr <- rand_forest(mtry = tune(), min_n = tune(), trees = 100) %>%
  set_engine("ranger", importance = "permutation") %>% 
  set_mode("classification")

nicotine_wflow_tr <-
  workflow() %>% 
  add_model(rf_model_tr) %>% 
  add_recipe(nicotine_recipe_tr)

rf_grid_tr <- grid_regular(
  mtry(range = c(2, 8)), 
  min_n(range = c(5, 20)),
  levels = 5
)

set.seed(123)
rf_res_tr <- tune_grid(
  nicotine_wflow_tr,
  resamples = rf_folds,
  grid = 10 
)
```

```{r}
rf_res_tr %>%
  collect_metrics() %>%
  filter(.metric == "accuracy") %>%
  select(mean, min_n, mtry) %>%
  pivot_longer(min_n:mtry,
    values_to = "value",
    names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "Accuracy")
```

```{r}
best_rf <- select_best(rf_res_tr, "accuracy")

final_rf <- finalize_workflow(
  nicotine_wflow_tr,
  best_rf
)

final_rf
```

```{r}
final_rf_fit <- fit(final_rf, train)
```

```{r}
final_rf_fit %>% pull_workflow_fit() %>% vip(geom = "point")
```

The five most impactful variables are sensation seeing (SS), openness to experience (Oscore), being from the UK, within the age range of 45-54 and impulsiveness. 

```{r}
trainpredrf <- predict(final_rf_fit, train)
head(trainpredrf)
```

```{r}
confusionMatrix(trainpredrf$.pred_class, train$Nicotine, 
                positive = "Yes")
```

```{r}
testpredrf <- predict(final_rf_fit, test)
head(testpredrf)

```

```{r}
confusionMatrix(testpredrf$.pred_class, test$Nicotine, 
                positive = "Yes")
```

Task 6 
The training set performed much better than the test set with a 92% accuracy to the test set's 71%.I would not say the results are consistent.


Task 7
I think a good predictive model could have nonprofit or commercial treatment implications.  Based on the categorical data certain demographics could be targeted.  I would not recommend this model for use in its current form.  There's a substantial difference in the training and testing accuracy results.  The test set only shows a 4% (71%) improvement over the No Information Rate of 67%. The discrepancy between the training and testing datasets are too great and further testing or refinement conducted.
  